---
title: "lab 9: How can we explain voter turnouts by age/race/...etc using data?"
author: "Erik Svenneby, Jorge Robelo, Mandy Karako, Ryan Bilkie, Thorbin Atwell"
date: "10/17/2019"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidyr)
library(dplyr)
library(readr)
library(devtools)
#website Erik found that has solid data: Electionlab.mit.edu/data
```

####Thorbin's question: How does wage and annual income correspond to voter turnout? 
```{r}
#thorbin's import: 
census_2011_data<-read_csv("~/Desktop/lab 9 data/Thorbin/census_2011_data")
```
For this lab, I took data from the 2012 census and Isolated variables including hours worked per day, university status and age. Using this and general census data regarding the 2012 election, I should be able to arrive at some possible variables to explain lack of younger voters. 

Hypothesis: 
I believe that younger people who are not in college, work a large amount of hours per week at lower salary leading to prolonged financial stress which significantly hampers motivation to vote. What I will use to confirm/deny this will be to compare yearly earnings and age which will demonstrate each age group's economic stress. 

A similar method will be employed on the college student sector. 
 
Since voting age is 18 in the US, and most students have graduated highschool by this age so by filtering age at and above 18 years, most non voting high school students will not be represented and most voting/non voting college students will be mostly represented. 


```{r}
voting_age_college<-census_2011_data %>% filter(AGEP>=18,SCHL>=18,!is.na(WKW))
voting_age_LF<-census_2011_data %>% filter(AGEP>=18,SCHL==16|17,!is.na(WKW))
ggplot(data=voting_age_college,mapping=aes(x=AGEP,y=WAGP))+geom_point()
```

From this plot it appears as though there is no correlation between wage/salary and age. This leads to the conclusion that salary age are not correlated. Just kidding, but what this data suggests is that salary does not play a role in voter turnout. 

Since I messed up and deleted the original data sets, what I would do differently if I had more time would be to first off select better variables namely hourly wage. I dont know how I missed this, but adding this into the dataframe would give a better metric to judge financial stress. Another thing I wanted to do was to bin an age range and examine the average income by age range, this would not only avoid the mess of the geom_point plot, but it would also be intuitave and easy to draw a conclusion off of.
Data to keep:
AGEP (age years) 
ST
CIT (citizen status , 5 is non citizen discard)
COW -class of worker
(b .Not in universe (less than 16 years old/NILF who last worked more .than 5 years ago or never worked)
1 .Employee of a private for-profit company or .business, or of an individual, for wages, .salary, or commissions
2 .Employee of a private not-for-profit, .tax-exempt, or charitable organization
           3 .Local government employee (city, county, etc.)
           4 .State government employee
           5 .Federal government employee
           6 .Self-employed in own not incorporated
             .business, professional practice, or farm
           7 .Self-employed in own incorporated
.business, professional practice or farm
8 .Working without pay in family business or farm
9 .Unemployed and last worked 5 years ago or earlier or never
.worked)
HINS1/2 Insurance 1=yes
PAP (welfare dollar amt)
SCH school enrollment
(2 .Yes, public school or public college
3 .Yes, private school, private college, or home school)-use in conj with SCHG neglect all values other than 15/16
SCHL
SEX 1=male
WKW weeks worked in past 12 months: 
b .N/A (less than 16 years old/did not work
           b .during the past 12 months)
           1 .50 to 52 weeks
           2 .48 to 49 weeks
           3 .40 to 47 weeks
           4 .27 to 39 weeks
           5 .14 to 26 weeks
           6 .13 weeks or less
PERNP/PINCP total persons earnings
WAGP (salary dollar ammount)

