---
title: "Lab 3 STAT_2600"
author: "Erik Svenneby,Jorge Robelo,Mandy Karako, Ryan Bilkie,Thorbin Atwell-Donaghey"
date: "10/9/2019"
output: html_document
---
# Black Swans lab 3
## Smoking bad?
Thorbin's Question: Is there a clear relationship between gender regardless of tobacco use? 
```{r echo=FALSE}
library(tidyverse)
library(dplyr)
source("http://www.openintro.org/stat/data/cdc.R")

```
```{r echo=FALSE}
first_step<-mutate(cdc,genhlth=case_when(genhlth=="poor"~1,genhlth=="fair"~2,genhlth=="good"~3,genhlth=="very good"~4,genhlth=="excellent"~5))
second_step<-mutate(first_step,corr_health=log(abs(weight-wtdesire),exp(1)))
cdc_transf<-mutate(second_step,corr_health=case_when(abs(weight-wtdesire)==0~genhlth,weight-wtdesire!=0~corr_health),gender=case_when(gender=="m"~1,gender=="f"~0),smoke100=case_when(smoke100==TRUE~"100 cig.club",smoke100==0~"non-smoker"))
cdc_transf_male<-filter(cdc_transf,gender==1)
cdc_transf_fem<-filter(cdc_transf,gender==0)
## this section transforms cdc into a more useable data frame ##
```
```{r echo=FALSE}
ggplot(data=cdc_transf_male,mapping=aes(x=age,y=corr_health,color=smoke100))+geom_point()+geom_smooth(color="black")+facet_wrap(~smoke100)+ggtitle("A plot comparing male smokers to male non-smokers")+labs(y="Corrected health value",x= "age (years)")+theme(legend.title=element_blank())
ggplot(data=cdc_transf_fem,mapping=aes(x=age,y=corr_health,color=smoke100))+geom_point()+geom_smooth(color="black")+facet_wrap(~smoke100)+ggtitle("A plot comparing female smokers to female non-smokers")+labs(y="Corrected health value",x= "age (years)")+theme(legend.title=element_blank())
## this section generates the plots ##
```
<<<<<<< HEAD
=======
end Thorbin


jorge's Question: what the relationship between age general health and smoking 
```{r echo=FALSE}
ggplot(data = cdc) + 
  geom_smooth(mapping = aes(x = age, y = smoke100))
```
#this graph displays the relationship between age and smoking. 40 percent of people from ages between 0-25 smoke. 
#following the age of 30 there is a steady increase in the percentage of people that smoke. 
#this increase persists until approximately the age of 60 where it is followed by a sharp decline in the percentage of smokers. 
#the decline in smokers persists until the age of 90 where we see another sharp incline in the percentage of smokers. 

#my graph 1 
```{r echo=FALSE}
ggplot(data = cdc) + 
  geom_smooth(mapping = aes(x = age, y = smoke100,  color = genhlth))
```
#this graph displays the relationship between age and smoking. 40 percent of people from ages between 0-25 smoke. 
#following the age of 30 there is a steady increase in the percentage of people that smoke. 
#this increase persists approximately until the age of 60 where it is followed by a sharp decline in the percentage of smokers. 
#the decline in smokers persist until the age of 90 where we see another sharp incline in the percentage of smokers. 
'''
end Jorge R
>>>>>>> 9dcfdaf34e524090e3328aa5b74cbf620056347a
