---
title: "Lab 7 Black Swans"
author: "Erik Svenneby, Jorge Robelo, Mandy Karako, Ryan Bilkie, Thorbin Atwell"
date: "10/16/2019"
output: html_document
---

```{r}

library(tidyverse)
library(readxl)
library(dplyr)
#Tang <- read_excel("East Africa lake data.xlsx")#
#Thorbin's import#
Tang<-read_excel("/Users/thorbinatwell-donaghey/Desktop/DSFA19/East_Africa_lake_data.xlsx",sheet=6)

```
```{r}
#Thorbin's lake visualisation

tang <- Tang %>% select("Year AD...1":TEX86) %>% mutate(year1=`Year AD...1`%/%1,year2=`Year AD...4`%/%1,year=Year%/%1) #%/%1 gives year as a whole number#
tang_BSI<-select(tang,year1,`BSi (normalized)`)
tang_CHARC<-select(tang,year2,`Charcoal (normalized)`)
tang_TEX86<-select(tang,year,TEX86)
tang_interim <- full_join(tang_BSI,tang_CHARC,tang_TEX86,by=c("year1"="year2"))
tang_final <-full_join(tang_interim,tang_TEX86,by=c("year1"="year"))

#Thorbin's plot# 
 ggplot(data=tang,mapping=aes(x=year,y=`BSi (normalized)`))+geom_point()+xlab("year")+ylab("Biogenic silica")+ggtitle("deviation of biogenic silica soil concentration by time")

```
Clearly concentration cannot be negative, the negative values indicate negative deviation from some kind of unity. If BSi concentration is a proxy of rain fall then higher positive deviations from unity indicate heavier periods of rain fall. 

###End Thorbin###
What we need:

* a general overall question you want to answer using data (you may end up changing your question next week)

* an explanation of why that question is important or interesting and a deeper explanation of why those reasons are important or     interesting

* a reference to at least one dataset you plan on using to answer this overall question
who the “pretend” domain expert is (and why they think the overall question is important and interesting)

* each person's individual African Lake visualization.
