---
title: "Lab 8 Black Swans"
author: "Erik Svenneby, Jorge Robelo, Mandy Karako, Ryan Bilkie, Thorbin Atwell"
date: "10/16/2019"
output: html_document
---
What we need:

* a general overall question you want to answer using data (you may end up changing your question next week)

* an explanation of why that question is important or interesting and a deeper explanation of why those reasons are important or     interesting

* a reference to at least one dataset you plan on using to answer this overall question
who the “pretend” domain expert is (and why they think the overall question is important and interesting)

* each person's individual African Lake visualization.
### Team Question:
We have been hired by the Oprah Winfrey 2024 presidential campaign manager to analyze the question of why voter turnout is so low.
This question is important because, in order to give her the greatest chance of winning, Oprah must understand what drives all demographics to the polls and what we can do to maximize her chances of success. Further, a better understanding of voting patterns may help us understand how humans function on a deeper, psychological level, which in turn can help give us better predictions about our political, economic and social future. At least one dataset we will be using is from census.gov: https://www.census.gov/topics/public-sector/voting.html

### Backup Question: 
What are the major causal factors driving violent crime?
The mayor of Denver Michael Hancock is determined to end violent crime once and for all in the city of Denver. To that end, he has hired our team to perform statistical analysis on crime data throughout the country in order to better understand what drives violent crime. This question is important because every data scientist would like to live in a world of peace and safety. At least one dataset we will be using is: https://catalog.data.gov/dataset/violent-crime-property-crime-by-county-1975-to-present-f0443

```{r}

library(tidyverse)
library(readxl)
library(dplyr)
#Tang <- read_excel("East Africa lake data.xlsx")#
#Thorbin's import#
#Tang<-read_excel("/Users/thorbinatwell-donaghey/Desktop/DSFA19/East_Africa_lake_data.xlsx",sheet=6)#

```
```{r}
#Thorbin's lake visualisation

tang <- Tang %>% select("Year AD...1":TEX86) %>% mutate(year1=`Year AD...1`%/%1,year2=`Year AD...4`%/%1,year=Year%/%1) #%/%1 gives year as a whole number#
tang_BSI<-select(tang,year1,`BSi (normalized)`)
tang_CHARC<-select(tang,year2,`Charcoal (normalized)`)
tang_TEX86<-select(tang,year,TEX86)
tang_interim <- full_join(tang_BSI,tang_CHARC,tang_TEX86,by=c("year1"="year2"))
tang_final <-full_join(tang_interim,tang_TEX86,by=c("year1"="year"))

#Thorbin's plot# 
 ggplot(data=tang,mapping=aes(x=year,y=`BSi (normalized)`))+geom_point()+xlab("year")+ylab("Biogenic silica")+ggtitle("deviation of biogenic silica soil concentration by time")

```
Clearly concentration cannot be negative, the negative values indicate negative deviation from some kind of unity. If BSi concentration is a proxy of rain fall then higher positive deviations from unity indicate heavier periods of rain fall. 

###End Thorbin###

##Ryan's Naivisha visualization

I chose to plot the depth of lake Naivasha over time:

'''{r}
Naivasha <- read_excel("C:\\Users\\ryanb\\OneDrive\\Desktop\\Data Science 2019\\East Africa lake data.xlsx", sheet=5)

Year = col_date(format="%Y")

Naivasha$Year <- parse_date(as.character(Naivasha$Year), format = "%Y")

Naivasha <- rename(Naivasha,LakeDepth=`Lake Depth`)
ggplot(data = Naivasha)+geom_col(mapping =aes(x=Year, y = LakeDepth), color = "blue")+geom_line(mapping =aes(x=Year, y = LakeDepth), color = "green")+xlab("Time")+ylab("Depth")
```

```{r}
library(tidyverse)
library(readxl)
Naivasha <- read_excel("C:\\Users\\ryanb\\OneDrive\\Desktop\\Data Science 2019\\East Africa lake data.xlsx", sheet=5)

Year = col_date(format="%Y")

Naivasha$Year <- parse_date(as.character(Naivasha$Year), format = "%Y")
#Doesn't work because of the decimals in the years.

Naivasha <- rename(Naivasha,LakeDepth='Lake Depth')
ggplot(data = Naivasha, mapping =aes(x='Year', y = 'Lake Depth'))+geom_point()+xlab("Time")+ylab("Depth")
```


Mandy's Lake Data: Victoria Lake 
```{r}
library(tidyverse)
library(dplyr)
library(readxl)
```
Importing data Victoria sheet of East Africa Lake Data: 
```{r}
East_Africa_lake_data <- read_excel("~/Downloads/East Africa lake data.xlsx", 
    sheet = "Victoria")
View(East_Africa_lake_data)
```
Assigning data:
```{r}
Victoria <- East_Africa_lake_data
```
Rename age to year:
```{r}
Victoria <- rename(Victoria, Year = 'Age')
```
Filter out NA's:
```{r}
Victoria <- filter(Victoria, !is.na(Year))
```
Round all years to a whole number: 
```{r}
year.victoria <- select(Victoria, Year)
year.victoria <- round(year.victoria, 0)
SWD.victoria <- select(Victoria, 'SWD(%)')
Victoria <- cbind(year.victoria, SWD.victoria)
```
Arrange years in descending order:
```{r}
Victoria <- arrange(Victoria, desc(Year))
```
Change name of SWD(%):
```{r}
Victoria <- rename(Victoria, SWDpercentage = 'SWD(%)')
```
Final tidyed data: 
```{r}
View(Victoria)
```
Plot: 
```{r}
ggplot(data=Victoria, mapping=aes(x=SWDpercentage, y=Year))+
  geom_point()+
  geom_smooth()+
  ggtitle("Percentage of SWD's per Year")
```
##End Mandy##
